---
title: "Seed Lots Work Flow"
output: html_document
date: "2024-10-10"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE, message = FALSE)
```

<br>

## Packages
```{r,  echo=TRUE}
library(tidyverse)
library(readxl)
library(kableExtra)
library(ggrepel)
library(readr)

```

## Winter 2024 seedlots
#### Seelots are created in T3 by transfering 'Grain Weight g' plot data to seed lots 
#### Create seedlots from a Trial
###### Trial can be specific managament factor subset of a trial (doesn't seem to work, freezes up)
###### Can use accession, plot, or sublot
###### {accession_name}-{trial_name}-plot_number

Enter a template for naming each new seedlot. The following variables (surrounded by curly braces) can be used in the template and the value of that variable will be replaced.

trial_name
accession_name
plot_name
plot_number
rep
block 
index - an incrementing index number, starting with 1

#### {accession_name}-{trial_name}-PLOT_{plot_number}
This format work well, it is best to have the plot number identified so that the harvest bag can be easily located. 

```{r,  echo=FALSE}

library(readr)
winter_2024_accession_max_seedlots <- read_csv("data/winter_2024_accession_max_seedlots.csv")%>% 
  mutate(seedlot = str_c(germplasmName,"-",observationUnitName)) %>%
  mutate(seed_weight_g = `Grain weight - g|CO_350:0005123`) %>%
  mutate(plotNumber = as.numeric(str_sub(observationUnitName,48))) %>% 
  mutate(trial = str_sub(observationUnitName,1,41)) %>%
  select(observationUnitName,germplasmName,seed_weight_g,seedlot,plotNumber,trial)


library(readxl)
Cornell_WinterOatFounders_2024_Headrow_phenotypes <- read_excel("data/Cornell_WinterOatFounders_2024_Headrow_phenotypes.xlsx") %>% 
 mutate(seed_weight_g = `Grain weight - g|CO_350:0005123`) %>% 
  mutate(seedlot = str_c(str_sub(observationUnitName,1,38),"-",germplasmName,"-1")) %>%
  mutate(plotNumber = as.numeric(str_sub(observationUnitName,40))) %>% 
  mutate(trial = str_sub(observationUnitName,1,38)) %>%
  select(observationUnitName,germplasmName,seed_weight_g,seedlot,plotNumber,trial)


seed_lots <- Cornell_WinterOatFounders_2024_Headrow_phenotypes %>% 
  bind_rows(winter_2024_accession_max_seedlots) 
```



```{r,echo=FALSE}

winter_2024_accession_max_seedlots %>% 
  select(seedlot, germplasmName, seed_weight_g)


```
### Ithaca 2025 trial seed needed and used
When the trial name is loooooooong it gets cumbersome to use.  Let's try and keep it shorter in the future.



### Ithaca 2025 trial seed needed and used




```{r,echo=FALSE}

plot_2025 <- read_excel("data/planning.xlsx", 
                       sheet = "selections")  %>% 
  mutate(trial = "plot_2025") %>% 
  mutate(seed_needed_g = plots_to_plant * 40.5) %>% 
  select(trial,germplasmName, plots_to_plant, seedlot,seed_needed_g)

guardrows <- read_excel("data/planning.xlsx", 
                       sheet = "guardrows")%>%
  mutate(plots_to_plant =1) %>% 
  mutate(seed_needed_g = plots_to_plant * 40.5) %>% 
  mutate(seedlot = str_c(germplasmName,"-",observationUnitName)) %>% 
  mutate(seed_used = plots_to_plant * 40.5) %>% 
  mutate(trial = "guardrows") %>% 
  select(trial,germplasmName, plots_to_plant, seedlot,seed_needed_g)

plot_increase <- read_csv("output/ plot_increase _randomization.csv") %>% 
  mutate(plots_to_plant =1) %>% 
  mutate(seedlot = str_c(substr(observationUnitName, start = 1, stop = 38),"-",germplasmName,"-1")) %>% 
  mutate(seed_needed_g = plots_to_plant * 40.5) %>% 
  mutate(trial = "plot_increase") %>% 
  select(trial,germplasmName, plots_to_plant, seedlot,seed_needed_g)

headrow_increase <- read_csv("output/ headrow_increase _randomization.csv") %>% 
  mutate(plots_to_plant =2) %>% 
  mutate(seed_needed_g = plots_to_plant * 6) %>% 
  mutate(seedlot = str_c(substr(observationUnitName, start = 1, stop = 38),"-",germplasmName,"-1")) %>% 
  mutate(trial = "headrow_increase") %>% 
  select(trial,germplasmName, plots_to_plant, seedlot,seed_needed_g)

winter_2025_seed_used <- bind_rows(plot_2025,
          guardrows,
          plot_increase,
          headrow_increase) %>% 
  group_by(seedlot) %>% 
  summarise(seed_used_g = sum(seed_needed_g))





```

```{r,echo=TRUE}

winter_2025_seed_used

```