---
title: "Randomization to T3 Trial"
author: "Peter Hyde"
date: "2024-11-15"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE, message = FALSE)
```

```{r,  echo=FALSE}
library(tidyverse)
library(readxl)
library(readr)
library(kableExtra)
```

<br>

<br>

### Overview

1.  **Import the Randomization**

2.  **Import the Field Map**

3.  **Enter Field Level Metadata**

4.  **Calculate Field Layout Metadata**

5.  **Format Plot level Metadata from Randomization**

6.  **Combine Field Level Metadata and Plot Level Metadata**

7.  **Save and import to T3/Oat**

<br>

<br>

### 1. Import the Randomization

The essintial data is 'plotNo', 'block', 'accession' and 'intercrop'.
Name the dataframe randomization

```{r,  echo=TRUE}

randomization <- winterOatPea_2025_plot_HR <- read_csv("output/winterOatPea_2025_plot_HR.csv") %>% 
  select(plotNo,block,accession,intercrop,crop)

randomization
```

### 2. Import the Map

The first column is row, it identifies the row number, my format has
them decending. The other column names are the number of the columns.
The map will be used to determine the row and column of each plot.

```{r,  echo=TRUE}

map <- CU_2025_Ithaca_WOP_PLOT_layout <- read_excel("data/2025 Oat-Pea Maps 9.24.24.xlsx", 
                                             sheet = "plot_layout")

map
```

### 3. Assign the field metadata

trial_name to trial_type are needed

##### trail_name convention "mother-ship"*"year"*"location"(short)*"trial abbreviation"*"type"(PLOT or HR)

##### design_type options

Alpha

Augmented

CRD

genotype_data_project

greenhouse

Lattice

p-rep

RCBD

RRC treatment

there are more, look on T3.

##### location
Search proper for location name on T3/oat.

```{r,  echo=TRUE}

trial_name = "CU_2025_Ithaca_WOP_PLOT"
breeding_program = "Intercropping Cooperative"
location = "Ithaca, NY - Snyder"
year = "2025"
design_type = ""
description =  "2025 winter oat pea intercrop trial located in Ithaca"
trial_type = "phenotyping_trial"
plot_width = "1.26"
plot_length = ""
field_size = ""     #this will be calculated and added in
planting_date =  "2024-10-04"
harvest_date  = ""

```

The upload file can be made manually using these headers

```{r,  echo=TRUE}
t3_trial_upload_headers <- c("trial_name",	"breeding_program",	"location",	"year",	"design_type",	"description",	"trial_type", "plot_width", "plot_length",	"field_size",	"planting_date", "harvest_date",	"plot_name",	"accession_name", "plot_number", "block_number",	"is_a_control",	"rep_number",	"range_number",	"row_number",	"col_number", "seedlot_name",	"num_seed_per_plot",	"weight_gram_seed_per_plot",	"entry_number", "is_private")

t3_trial_upload_headers

```

### 4. Calculate Field Metadata

Two methods for calculating total number of plots make sure they match!

```{r,  echo=TRUE}
length(randomization$plotNo) # how many plots are there?
max(randomization$plotNo) # how many plots are there?

if_else(length(randomization$plotNo) == max(randomization$plotNo), "ALL GOOD" , "DANGER")

total_plots  <- length(randomization$plotNo)

total_plots


# row_number, col_number and field size in hectares

randomization_row_col  <-  map %>% 
  pivot_longer(!row, names_to = "column", values_to = "plot") %>% 
  mutate(row_number = row) %>% 
  mutate(col_number = as.numeric(column)) %>% 
  mutate(plot_number = plot) %>% 
  select(plot_number,	row_number,col_number) %>% 
  filter(!is.na(plot_number)) %>% 
  arrange(plot_number)

row_number <- randomization_row_col$row_number
col_number <- randomization_row_col$col_number


field_size = ((max(randomization_row_col$row_number)) * 5.5) * 
            ((max(randomization_row_col$col_number)) * 1.5 ) / 10000

field_size  ### field size in hectares
```

### 5. Create two dataframes with trial level and plot level metadata

#### Trial level meta

```{r}

trial_level_headers <- c("trial_name","breeding_program","location","year","design_type","description","trial_type","plot_width","plot_length","field_size","planting_date","harvest_date")


trial_meta <- bind_cols(trial_name,  #keep in this order, to match t3 template
          breeding_program,
          location,
          year,
          design_type,
          description,
          trial_type,
          plot_width,
          plot_length,
          field_size,
          planting_date,
          harvest_date)

colnames(trial_meta) <- trial_level_headers

# duplicate the trial level meta to match the number of plots
trial_meta <- trial_meta  %>%  slice(rep(1:n(), each = total_plots)) 

trial_meta %>% 
  kable(align = "l") %>% 
  kable_classic(full_width=F,position = "left")%>% 
  scroll_box(width = "900px", height = "300px")

```

#### Plot level metadata

```{r}

plot_level_headers <- c("plot_name","accession_name","plot_number","block_number",	"is_a_control","rep_number","range_number","row_number","col_number","seedlot_name","num_seed_per_plot","weight_gram_seed_per_plot","entry_number",	"is_private")

randomization

number_of_blocks <- length(unique(randomization$block))

```

#### Quality Double Check

```{r}

# check if the number of plots in each block are equal

plots_per_block <- unique(table(randomization$block))

# if the number of plots per block are not equal "length(plots_per_block)" will be greater than 1
stopifnot(length(unique(table(randomization$block)))==1)  

# check that the plot and block numbers match up
stopifnot(total_plots == number_of_blocks*plots_per_block)


table(randomization$block)
total_plots
number_of_blocks
plots_per_block
```

#### Pea Accession

List all the pea accessions

```{r}

pea_accessions <- randomization %>% 
  filter(crop == "Pea") %>%
  select(accession) %>% 
  unique()

pea_accessions

```

### Data from randomization

Take the randomization and convert the sequential plot numbers to 3
digit plot numbers starting with the block followed by the sequential
number of plots in the block. Monocrop pea plots accession is set to
NO_OATS_PLANTED. Plots with the accession listed as Blaze (a pea) will
identified by a 1 in a column headed Blaze. The column 'Blaze' will be
converted by T3 into a management factor with the name 'Blaze' used as
written in the column heading.

```{r}

trial <- randomization %>% 

  # change the plot numbers from sequential to 3 digit with block and plots per block
  mutate(plot_number = (rep(seq(1:plots_per_block), times = number_of_blocks)) +   
           (rep(seq(1:number_of_blocks), each = plots_per_block)*100)) %>% 

  # change pea accessions to NO_OATS_PLANTED which is a valid t3 oat accession
  mutate(accession_name = if_else(crop == "Pea", "NO_OATS_PLANTED",accession)) %>% 

  # need to add in for each pea accession.  This will be used to create the pea management factor 
  mutate(Blaze = if_else(accession == "Blaze"|intercrop == "intercrop", "1",""))  

trial
  
```

Add in T3/oat plot level meta data. If no data is available use "" for
blank.

```{r}
trial <- trial %>% 
  
  mutate(plot_name = str_c(trial_name,"_",plot_number)) %>% 
  
  mutate(block_number  = as.numeric(str_remove(randomization$block, "Block" ))) %>% 
  mutate(is_a_control = "") %>% 
  
  mutate(rep_number = "") %>% 
  
  mutate(range_number = "") %>% 
  
  mutate(row_number = row_number) %>%
  
  mutate(col_number = col_number) %>% 
  mutate(seedlot_name = "") %>% 
  mutate(num_seed_per_plot = "") %>% 
  mutate(weight_gram_seed_per_plot = "") %>% 	
  mutate(entry_number = "") %>% 
  mutate(is_private = "")
  

trial <- trial %>% 
  select(plot_name,
          accession_name,
          plot_number,
          block_number,
          is_a_control, 
          rep_number,               
          range_number,
          row_number,     
          col_number,             
          seedlot_name,         
          num_seed_per_plot,      
          weight_gram_seed_per_plot,
          entry_number,
          is_private,
          Blaze)   # need to add in peas at the end, plots with peas get a 1 others are empty. This will be converted to management factor upon upload.


trial %>% 
  kable(align = "l") %>% 
  kable_classic(full_width=F,position = "left")%>% 
  scroll_box(width = "900px", height = "300px")


```

### 6. Combine Field Level Metadata and Plot Level Metadata

Merge the plot level data with the trial level data.

```{r}

full_trial <- bind_cols( trial_meta ,trial)

full_trial%>% 
  kable(align = "l") %>% 
  kable_classic(full_width=T,position = "left")%>% 
  scroll_box(width = "900px", height = "300px")

```

### 7. Save and import in T3/oat

Save the trial The csv will need to be converted to excel for upload Use
multi-trial upload in T3/oat.

```{r}

## write.csv(full_trial,str_c("output/",trial_name,"_t3_upload.csv"), row.names = FALSE)


```
