---
title: "Randomization"
author: "Peter Hyde"
date: "2024-11-14"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE, message = FALSE)
```

```{r,  echo=FALSE}
library(readxl)
library(AlgDesign)
library(tidyverse)
library(readxl)
```

<br>

<br>

### Overview of general steps

1.  **Import a list of accessions that are to be used in trial**

2.  **Select the list and name accessions**

3.  **Set blocks, and accessions per block**

4.  **Randomize**

5.  **Format for printing envelopes**

<br>

<br>

### 1. Import a list of T3/oat accessions

```{r,  echo=FALSE}

input_file <- read_excel("data/2025_SOAP_Entries.xlsx") %>% 
  mutate(accession = Name)

input_file
```

<br>

### 2. Create accession list

Make sure the heading is 'accessions'

```{r,  echo=TRUE}
accessions <- input_file %>% 
  mutate(accession = Name) %>% # rename as accession
  select(accession) %>% 
  distinct()  # ensure accession list has no duplicates
```

<br>


### 3. Set the number of blocks and accessions per block

In this example there are 455 accessions and all accessions are in each
block. There are 2 blocks. It is a CRBD.

```{r,  echo=TRUE}

nBlks <- 2   # set the block number

nAccessions <- nrow(accessions)  # how many accessions in the plan

nPlotsPerBlk <- nAccessions # set the number of plots per block
```

Create a dataframe with the accessions duplicated for each block
```{r,  echo=TRUE}
blocks <- dplyr::tibble( 
  accession=rep(accessions$accession, nBlks),
  block= rep(1:nBlks,  each = nPlotsPerBlk ) )

blocks
```

<br>


### 4. Randomize
set.seed to make the randomization reproducible 

```{r,  echo=TRUE}

floor(runif(1, 10000, 1000000))  # run to get a random number for set.seed

set.seed(892015) 
randomization  <- blocks %>% 
  left_join(input_file, join_by(accession)) %>% ### add in metadata from input file
  mutate("random" = runif(1:nrow(blocks))) %>% 
  arrange(block,random) %>%   # order by random numbers
  mutate(plot = seq(1:nrow(blocks))) %>%  # add in sequential plot numbers
  relocate(block,plot,accession)


randomization

```

```{r,  echo=TRUE}
# write.csv(randomization,"output/SOAP_randomization.csv",row.names = FALSE)

```

### 5. Formatted output for printing labels

arranged by entry and save as a .csv for use on enevelope printer

```{r,  echo=TRUE}

randomization %>% 
  arrange(Entry) 

# write.csv("output/SOAP_labels.csv",row.names = FALSE)

```
