---
title: "Seedlots and Seedpacking Workflow"
output: html_document
date: "2024-10-10"
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE, message = FALSE)
```

```{r,  echo=FALSE}
library(tidyverse)
library(readxl)
library(kableExtra)
library(ggrepel)
library(readr)

```

## 

### Overview of general steps

1.  **Create seedlot**

2.  **Pick accessions to plant for a new trial**

3.  **Join new trial accessions to seedlot accessions**

4.  **Pull bulk seed packs**

5.  **Pack plot envelops**

<br>

<br>

#### 1. Create seedlots from a Trial

Seedlots are created in T3 by transferring 'Grain Weight g' plot data to
seed lots. When creating the seedlot a template is used for naming the
seedlot. Have the **source plot** number identified so that the harvest
bags can be easily located.

<br>

This format work well.

##### {accession_name}-{trial_name}-PLOT\_{plot_number}

-   If the trial name is long it becomes a problem printing the seedlot
    name on envelopes or labels. In the future it might be best to keep
    the trial names shorter or use and abbreviation when creating the
    seedlot names. {accession_name}-trial_abbr.-PLOT\_{plot_number}

<br>

Sources for seed lots can be accession, plot, or subplot

-   Functionally, we source the seed by ***plot*** and ***greatest seed
    weight*** for each given accession. These seed bags are separated
    out and used to fill seed packs. That way we are dealing with fewer
    bags. For example the 2024 winter trial had 378 plots but 96
    accession. We worked with 96 bags.
-   Sub-setting the seedlot sorce by management factor looks to be
    available but doesn't seem to work, it freezes up. This could be
    useful in if we wanted to select only mono-crop plots for seed
    purity purposes.

<br>

The following variables can also be used in the template

\| trial_name

\| accession_name

\| plot_name

\| plot_number

\| rep block

\| index - an incrementing index number, starting with 1

<br>

<br>

### Example Seedlots

These are the seedlots created from
`Cornell_WinterOatPeaIntercrop_2024_Ithaca` and
`Cornell_WinterOatFounders_2024_Headrow` They include only the plot with
the maximum seed weight for each accession.

```{r,  echo=FALSE}

library(readr)
winter_2024_accession_max_seedlots <- read_csv("data/winter_2024_accession_max_seedlots.csv")%>% 
  mutate(seedlot = str_c(germplasmName,"-",observationUnitName)) %>%
  mutate(seed_weight_g = `Grain weight - g|CO_350:0005123`) %>%
  mutate(plotNumber = as.numeric(str_sub(observationUnitName,48))) %>% 
  mutate(trial = str_sub(observationUnitName,1,41)) %>%
  select(observationUnitName,germplasmName,seed_weight_g,seedlot,plotNumber,trial)


library(readxl)
Cornell_WinterOatFounders_2024_Headrow_phenotypes <- read_excel("data/Cornell_WinterOatFounders_2024_Headrow_phenotypes.xlsx") %>% 
 mutate(seed_weight_g = `Grain weight - g|CO_350:0005123`) %>% 
  mutate(seedlot = str_c(str_sub(observationUnitName,1,38),"-",germplasmName,"-1")) %>%
  mutate(plotNumber = as.numeric(str_sub(observationUnitName,40))) %>% 
  mutate(trial = str_sub(observationUnitName,1,38)) %>%
  select(observationUnitName,germplasmName,seed_weight_g,seedlot,plotNumber,trial)


seed_lots <- winter_2024_accession_max_seedlots %>% 
  bind_rows(Cornell_WinterOatFounders_2024_Headrow_phenotypes) 
```

```{r,echo=FALSE}

seed_lots %>% 
  select(seedlot, germplasmName, seed_weight_g)%>% 
  kable(align = "l") %>% 
  kable_classic(full_width=F,position = "left")%>% 
  scroll_box(width = "900px", height = "300px")



```

<br>

<br>

### 2. **Pick accessions to plant for a new trial**

```{r,echo=FALSE}

plot_2025 <- read_excel("data/planning.xlsx", 
                       sheet = "selections")  %>% 
  mutate(trial = "plot_2025") %>% 
  mutate(seed_needed_g = plots_to_plant * 40.5) %>% 
  select(trial,germplasmName, plots_to_plant, seedlot,seed_needed_g)

guardrows <- read_excel("data/planning.xlsx", 
                       sheet = "guardrows")%>%
  mutate(plots_to_plant =1) %>% 
  mutate(seed_needed_g = plots_to_plant * 40.5) %>% 
  mutate(seedlot = str_c(germplasmName,"-",observationUnitName)) %>% 
  mutate(seed_used = plots_to_plant * 40.5) %>% 
  mutate(trial = "guardrows") %>% 
  select(trial,germplasmName, plots_to_plant, seedlot,seed_needed_g)

plot_increase <- read_csv("output/ plot_increase _randomization.csv") %>% 
  mutate(plots_to_plant =1) %>% 
  mutate(seedlot = str_c(substr(observationUnitName, start = 1, stop = 38),"-",germplasmName,"-1")) %>% 
  mutate(seed_needed_g = plots_to_plant * 40.5) %>% 
  mutate(trial = "plot_increase") %>% 
  select(trial,germplasmName, plots_to_plant, seedlot,seed_needed_g)

headrow_increase <- read_csv("output/ headrow_increase _randomization.csv") %>% 
  mutate(plots_to_plant =2) %>% 
  mutate(seed_needed_g = plots_to_plant * 6) %>% 
  mutate(seedlot = str_c(substr(observationUnitName, start = 1, stop = 38),"-",germplasmName,"-1")) %>% 
  mutate(trial = "headrow_increase") %>% 
  select(trial,germplasmName, plots_to_plant, seedlot,seed_needed_g)

winter_2025_seed_used <- bind_rows(plot_2025,
          guardrows,
          plot_increase,
          headrow_increase) %>% 
  group_by(seedlot) %>% 
  summarise(seed_used_g = sum(seed_needed_g))





```

New trail is planned (another script). The output of the planning is a
table with `new_trial`, `germplasmName`, `plots_to_plant`, and
`source_trial` (or abbreviation to match seedlot name). The source trial
name is used to create the seed lot name that will be used to match with
the seed lot.

Seed needed is calculated by multiplying the number of plots by the seed
per plot.

```{r, echo=FALSE}
new_trial <- read_excel("data/planning.xlsx", 
                       sheet = "selections")  %>% 
  mutate(new_trial = "plot_2025") %>% 
  mutate(seed_needed_g = plots_to_plant * 40.5) %>%
  mutate(source_trial = str_c(substr(observationUnitName, start = 1, stop = 41))) %>% 
  mutate(seedlot = str_c(germplasmName,"-",observationUnitName)) %>% 
  select(new_trial, germplasmName, source_trial, plots_to_plant, seed_needed_g) 

new_trial %>% 
  kable(align = "l") %>% 
  kable_classic(full_width=F,position = "left")%>% 
  scroll_box(width = "900px", height = "300px")

```

### 3. Join new trial accessions to seedlot accessions

**New Trial**

Create `seedlot_no_plot` by concatenating `germplasm` and
`source_trail`. `Seedlot_no_plot` should match the seedlot names
excluding the plot number.

```{r,echo=TRUE}

new_trial %>% 
  mutate(seedlot_no_plot = str_c(germplasmName,"-", source_trial,"-PLOT_" )) %>%  kable(align = "l") %>% 
  kable_classic(full_width=F,position = "left")%>% 
  scroll_box(width = "900px", height = "300px") 
  
```

<br>

**Seedlot**

Create `seedlot_no_plot` from seed lot by removing plot number

```{r,echo=TRUE}

seed_lots %>% 
  select(seedlot, germplasmName, seed_weight_g) %>% 
  mutate(seedlot_no_plot = gsub("(PLOT_).*","\\1",seed_lots$seedlot))%>%  
  kable(align = "l") %>% 
  kable_classic(full_width=F,position = "left")%>% 
  scroll_box(width = "900px", height = "300px") 
  

```

<br>

**Join the New Trial with Seedlot**

```{r, echo=FALSE}

bulk_seed <- new_trial %>% 
  mutate(seedlot_no_plot = str_c(germplasmName,"-", source_trial,"-PLOT_" )) %>% 
  left_join((seed_lots %>% 
    select(seedlot, germplasmName, seed_weight_g) %>% 
    mutate(seedlot_no_plot = gsub("(PLOT_).*","\\1",seed_lots$seedlot) )),by=join_by(seedlot_no_plot)) %>% 
  select(new_trial,source_trial,plots_to_plant,seed_needed_g,seedlot,seed_weight_g) %>% 
  mutate(seed_left = seed_weight_g - seed_needed_g) 

bulk_seed %>%  kable(align = "l") %>% 
  kable_classic(full_width=F,position = "left")%>% 
  scroll_box(width = "900px", height = "300px") 
```

<br>

Check if there is enough seed available. Any accessions with less than
10 grams will be listed

```{r}

bulk_seed %>% 
  filter(seed_left < 10)


```

<br>

<br>

### 4. Bulk seed labels

Bulk seed bags contain the seed for all plots of a given accession and
will be divided into plot envelopes in the next step. Labels need to
have **source plot number**. Sort them by Source_trial and plot number

```{r}

bulkseed_2 <- bulk_seed %>% 
  select(seedlot, seed_needed_g, plots_to_plant,source_trial) %>% 
  mutate(source_plot = as.numeric(gsub(".*PLOT_","",.$seedlot))) %>% 
  mutate(accession = gsub("-.*","",.$seedlot)) %>% 
  mutate(source_trial = gsub("-PLOT_.*","",.$seedlot)) %>%
  mutate(source_trial = gsub(".*-","",.$source_trial)) %>%
  arrange(source_trial,source_plot)  


bulkseed_2 %>% 
  kable(align = "l") %>% 
  kable_classic(full_width=F,position = "left")%>% 
  scroll_box(width = "900px", height = "300px")

```

<br>

Make a label using mail merge in excel and word.

Example:

**NF01401B**

Cornell_WinterOatPeaIntercrop_2024_Ithaca-PLOT_1

162 grams for 4 plots

-   It would be great to do this in T3.

<br>

<br>

### 5. Plot seed labels

The envelope printer uses a .csv file and **must have headers in this
order** `source`, `plotNo`, `oatName`, `peaName.` Arrange by source plot
so all plots for a given accession are together for seed separating from
bulk bag. The Oat-Pea template is on the printer
computer/desktop/Jannink.

<br>

**Full Randomization (spripts/winter_2025_randomization_HR_and_plot.R)**

```{r, echo=FALSE}

winterOatPea_2025_plot_HR <- read_csv("output/winterOatPea_2025_plot_HR.csv") %>% 
  select(plotNo:crop_2)

winterOatPea_2025_plot_HR %>% 
  kable(align = "l") %>% 
  kable_classic(full_width=F,position = "left")%>% 
  scroll_box(width = "900px", height = "300px")


```

**Join Randomization with Bulk_Seed by accession**

```{r,echo=FALSE}

winterOatPea_2025_plot_HR %>% 
  left_join(bulkseed_2, by= join_by(accession)) %>% 
  mutate(source_trial = gsub("-PLOT_.*","",.$seedlot)) %>% 
  mutate(source_trial = gsub(".*-","",.$source_trial)) %>% 
  mutate(peaName = if_else(crop_2=="oat-pea","Blaze","na")) %>% 
  mutate(oatName = accession) %>% 
  mutate(PlotNo = plotNo) %>% 
  mutate(source = seedlot) %>% 
  arrange(source_plot) %>% 
  select(source,plotNo,oatName,peaName)

```

Save as .csv without row.names *write.csv(
"output/winter_2025_envelopes.csv", row.names = FALSE)*

Put it on a thumb drive and take it to the printer.
